---
title: "Anthropogenic Noise Impacts of Frog Vocalization at Anastasia State Park and the GTM Research Reserve"
author: "Emma Wilkinson, Dakota Chenoweth, Trisha McCaul"
date: "11/14/2021"
output: html_document
---

```{r, include = FALSE}
library(tidyverse)
library(dplyr)
library(ggpubr)
library(PairedData)
library(rstatix)

asp <- read.csv("/Volumes/SeaGate Wilkinson/Capstone2021/captsone_2021_frogs/data_raw/ASP_data.csv", header = TRUE)
gtm <- read.csv("/Volumes/SeaGate Wilkinson/Capstone2021/captsone_2021_frogs/data_raw/GTM_data.csv", header = TRUE)
```

## Species composition charts
### Anastasia State Park (ASP)

```{r, echo = FALSE}

#ASP

species <- asp %>% count(speciesID)
species

cols <- c("#617335", "#4F8A8F", "#B0BC41", "#BBD3F4", "#A6A6A6")

ggplot(data = species, aes(x = speciesID, y = n)) +
    geom_bar(stat = "identity", color = cols, fill = cols) +
    labs(size = 26,
         title = "ASP Identified Species Composition",
         x = "Species ID",
         y = "Count") +
    geom_text(aes(label = n),
              vjust = -0.3,
              size = 3.5)

```

### Guana Tolomato Matanzas National Estuarine Research Reserve (GTM)
```{r, echo = FALSE}

#GTM

species <- gtm %>% count(SpeciesID)
species

cols <- c("#617335", "#B0BC41", "#A6A6A6")

ggplot(data = species, aes(x = SpeciesID, y = n)) +
    geom_bar(stat = "identity", color = cols, fill = cols) +
    labs(size = 26,
         title = "GTM Identified Species Composition",
         x = "Species ID",
         y = "Count") +
    geom_text(aes(label = n),
              vjust = -0.3,
              size = 3.5)

```


## One Way Repeated Measures ANOVA

ANOVA compares differences in population means between multiple related samples.

H0: μ before = μ during = μ after

H1: at least one μ is different

```{r, echo = FALSE}

#summarize and visualize data
summarydf <- asp %>%
    gather(key = "treatment", value = "call_rate", RateBefore, RateDuring, RateAfter) %>%
    convert_as_factor(treatment) #prepare a df with a column called "treatment" that shows before/during/after

#add an ID column to df bc the ANOVA test needs it
summarydf$ID <- seq(1: 282) #282 is the dim of summarydf

summarydf <- summarydf %>% convert_as_factor(ID) #convert ID into a factor type

summarydf %>%
    group_by(treatment) %>%
    get_summary_stats(call_rate, type = "mean_sd") #get mean call rate and sd of treatments

```


```{r, echo = FALSE}

bxp <- ggboxplot(summarydf,
                 x = "treatment",
                 y = "call_rate",
                 xlab = "Treatment",
                 ylab = "Call rate",
                 main = "Call Rates Before, During, and After Noise Events at ASP",
                 order = c("RateBefore", "RateDuring", "RateAfter"),
                 font.label = list(size = 26, color = "black"),
                 ggtheme = theme_gray()) # create boxplot of treatments
bxp

```


```{r, echo = FALSE}
### Check ANOVA assumptions

#identify extreme outliers
# extremeo <- summarydf %>%
#     group_by(treatment) %>%
#     identify_outliers(call_rate) #no extreme outliers

#assess normality
# ggqqplot(summarydf, "call_rate", facet.by = "treatment")
#if sample size is greater than 40 normality can be assumed

```

```{r}
### Run ANOVA

anova.1 <- aov(call_rate ~ treatment, data = summarydf)

summary(anova.1)
```

The P-value was > a = 0.05, so we fail to reject the possibility that all three population means could be equal.

## Two Sample T-Test

T-test compares the differences in means between two unrelated samples

H0: μ ASP = μ GTM

H1: μ ASP ≠ μ GTM

```{r, echo = FALSE}

#We predict that the mean population call rate of species at asp will be different than the mean population call rate of species at the gtm
#H0:m asp = m gtm
#Ha:m asp != m gtm

#summarize and visualize data
ttestdf <- read.csv("/Volumes/SeaGate Wilkinson/Capstone2021/captsone_2021_frogs/data_raw/2_samp_ttest_df.csv", header = TRUE)

#summary stats
group_by(ttestdf, Site) %>%
    summarise(
        count = n(),
        mean = mean(CallRate, na.rm = TRUE),
        sd = sd(CallRate, na.rm = TRUE)
    )


```


```{r, echo = FALSE}

#plot
bxp <- ggboxplot(ttestdf, x = "Site", y = "CallRate", 
          color = "Site", palette = c("#000000", "#000000"),
          ylab = "Call Rate", xlab = "Site",
          main = "ASP has Larger Distribution of Call Rate",
          font.label = list(size = 26, color = "black"),
          ggtheme = theme_gray())
bxp <- ggpar(bxp, legend = "none")
bxp

```

#### check t-test assumptions

```{r, echo = FALSE}

### check assumptions
# are the populations independent?
    # yes. different locations
# do the data follow normal distribution?
# ggqqplot(ttestdf, "CallRate", facet.by = "Site")
#normality can be assumed in a sample size greater than 40
#do the populations have the same variance?
ftest <- var.test(CallRate ~ Site, data = ttestdf)
ftest
# the variances are significantly different. We need to use Welch ttest

```
the variances are significantly different, so we use the Welch t-test to account for that.

#### compute Welch two sample t-test
```{r}

### compute t test
ttest <- t.test(CallRate ~ Site, data = ttestdf, alternative = c("two.sided"), var.equal = FALSE) #var.equal = FALSE b/c population variances != -- this runs a Welch ttest to make up for it
ttest
#pval = 0.72

```

The P-value > a= 0.05, so we fail to reject the possibility that the population mean call rate of ASP = the population mean call rate of the GTM.